Surface correlation functions calculations: improving methodology for discrete images with novel exact continuous approach

### 2023-03-02 paper.tex
Very important commentary
Add bibliography
Add a note about isotropic medium
Introduction

\author{Vasily Postnicov\textsuperscript{1}}
\author{Marina Karsanina\textsuperscript{1}}
\author{Efim Lavrukhin\textsuperscript{1,2}}
\author{Dina Gafurova\textsuperscript{3}}
\author{Aleksey Khlyupin\textsuperscript{1}}
\author{Kirill Gerke\textsuperscript{1}}
\email{kg@ifz.ru}

\affiliation{\textsuperscript{1}Schmidt Institute of Physics of the Earth of
  Russian Academy of Sciences, Moscow, 107031, Russia}
\affiliation{\textsuperscript{2}Computational Mathematics and Cybernetics,
  Lomonosov Moscow State University, Moscow, 119991, Russia}
\affiliation{\textsuperscript{3}Oil and Gas Research Institute Russian Academy
  of Sciences (OGRI RAS) 3, Gubkina st., Moscow, 119333, Russian Federation}

\title{We did an amazing thing}

@@

\section{Introduction}
\label{sec:intro}
AAA

@@

Now define two disjoint sets in a homogeneous porous medium: a set of solid
phase $S$ and a set of void phase $V$. Surface-surface correlation function
$F_{ss}$ and surface-void correlation function $F_{sv}$ are defined as follows:
\begin{align}
  F_{ss}(\bm{r}) &= \langle |\nabla \chi_S(\bm{x})| |\nabla \chi_S(\bm{x} +
  \bm{r})| \rangle \label{eq:fss} \\
  F_{sv}(\bm{r}) &= \langle |\nabla \chi_S(\bm{x})| \chi_V(\bm{x} +
  \bm{r}) \rangle \label{eq:fsv}
\end{align}
Here angles $\langle \dots \rangle$ denote ensemble average. For isotropic media
vector values $\bm{x}$ and $\bm{r}$ can be replaced with scalars, but in this
paper we will consider anisotropic media as well as isotropic.

More detailed information on surface correlation functions can be found in a
book of Salvatore Torquato\cite{Torquato_book}.

@@

@@

\subsection{Surface-surface function in two-dimensional case}
\label{sec:fss-2d}
\begin{figure}
  \centering
  \includegraphics[width=0.8\linewidth]{images/Fss.png}
  \caption[]{Illustration of how $F_{ss}(\bm{r})$ is calculated. The interface
    between solid and void phases is a solid black curve. This interface
    translated by the vector $\bm{r}$ is show as a red solid curve. Support of
    $f_\Gamma(\bm{x}; \Delta)$ is a dashed ring. Shaded regions contribute to
    the surface-surface function.}
  \label{fig:Fss-explained}
\end{figure}
Let $S$ be some set in $\mathbb{R}^2$ which has a smooth boundary $\Gamma$ in
the sense that the boundary has a tangent line almost at every point. Now we
introduce a function $f_\Gamma(\bm{x}; \Delta)$:
\begin{equation*}
  f_\Gamma(\bm{x}; \Delta) = \left\{
  \begin{array}{ll}
    1/\Delta & \quad \rho(\bm{x}, \Gamma) < \Delta \\
    0 & \quad \text{otherwise}
  \end{array}
  \right.
\end{equation*}
Here $\rho(\bm{x}, \Gamma)$ means a closest distance from point $\bm{x}$ to the
boundary $\Gamma$.

@@

It's useful to introduce an imaginary unit $\varepsilon$ and write a dual number
$(x, y)$ as $x + y\varepsilon$. As you can see from multiplication law,
$1\cdot \varepsilon = \varepsilon$ and $\varepsilon^2 = 0$, hence $D$ is not an
integral domain and division is not defined for all dual numbers with exception
of $0$. Namely it's defined only if divisor has non-zero real part:
\begin{align*}
  \frac{a+b\varepsilon}{c+d\varepsilon} &=
  \frac{(a+b\varepsilon)(c-d\varepsilon)}{(c+d\varepsilon)(c-d\varepsilon)} \\
  &= \frac{ac-ad\varepsilon+bc\varepsilon}{c^2} \\
  &= \frac{a}{c} + \frac{bc-ad}{c^2}\varepsilon
\end{align*}

@@

>>>>

\author{Vasily Postnicov\textsuperscript{1}}
\author{Marina V. Karsanina\textsuperscript{1}}
\author{Aleksey Khlyupin\textsuperscript{1}}
\author{Kirill M. Gerke\textsuperscript{1}}
\email{kg@ifz.ru}

\affiliation{\textsuperscript{1}Schmidt Institute of Physics of the Earth of
  Russian Academy of Sciences, Moscow, 107031, Russia}
%\affiliation{\textsuperscript{2}Computational Mathematics and Cybernetics,
%  Lomonosov Moscow State University, Moscow, 119991, Russia}
%\affiliation{\textsuperscript{3}Oil and Gas Research Institute Russian Academy
%  of Sciences (OGRI RAS) 3, Gubkina st., Moscow, 119333, Russian Federation}

\title{Surface correlation functions calculations: improving methodology for discrete images with novel exact continuous approach}

@@

\section{Introduction}
\label{sec:intro}
Correlation functions (CFs) to describe structures arose from scattering experiments \cite{debye1957scattering} and
found their way into numerous disciplines and applications: from material sciences \cite{Cecen,chen2022quantifying}, to rock
\cite{ledesma2018effect} and soil physics \cite{Euras2012,PLoS_ONE,KarsaninaEJSS}, cosmology \cite{TakadaJain} and food
engineering \cite{Derossi2019}. In particular, correlation functions as computed from structural images were used for:
\begin{enumerate}
  \item characterization of morphology \cite{tensorPRE} and representativeness via
  correlation lengths \cite{vcapek2011transport,thovert2011grain} at which CFs reach plateau;
  \item comparison of different structures \cite{jiao2014chawla,EPL1,REVpaper}, usually by computing L-2 norm difference between CFs for two or more images,
  \item compressing structural information \cite{SciRep1,Havelka} by representing huge datasets in the form of a limited set of CFs that can be further squeezed into parameters for a number of basis functions \cite{jiao_basis,KarsaninaEJSS},
  \item describing structural dynamics and evolution \cite{chen2015dynamic,PLoS_ONE,xu2022correlation},
  \item extracting features for deep learning \cite{Miao2017,kamrava2020linking,roding2020predicting,KarsaninaEJSS},
  \item performing stochastic reconstructions \cite{Adler_recon,Y-T,EPL2,tahmasebiPRL,karsaninaPRL} based on limited input data such as, for example, CFs obtained from 2D cuts,
  \item fusing multi-scale images \cite{SciRep1,chen2016stochastic,Geoderma2018} into a single representation,
  \item evaluation of image replica quality, e.g., stochastic reconstructions \cite{deeep_papers_or_somethingelse},
  \item assessing stationarity of the structure at hand \cite{REVpaper}.
\end{enumerate}

In majority, if not all, cases mentioned above, surface functions - namely, $F_{ss}$ and $F_{sv}$ - were underrepresented due to our inability to compute them in a robust way. While known for a long time \cite{theory_in_MT}, only somewhat crude approximate approaches \cite{crude_approx_inMT} were developed until Ma and Torquato laid foundation for such a robust computational methodology \cite{ma2018SS}. They proposed an elegant solution for infinite resolution Gaussian random fields, or "continouos" approach. They also assumed that the same approach can be readily applied to discrete experimental images if grey-scale values are thresholded in a similar way to continuious Gaussian field. Samarin et al. \cite{our_SS_paper} argued that experimental images needs to be segmented first using current cutting edge techniques \cite{NNseg} and proposed "discrete" approach. The latter allowed to obtain surface functions closer to analytical solutions for overlapping disks/spheres and to supress noise, in particular for $F_{ss}$. The noise in Ma and Torquato method resulted from $1/cos(\theta)$ computations during random line segment placement - this leaves a room for improvements. While Samarin et al. obtained better convergence with analytical solution, this was demostrated only for overlapping disks image of limited size (that can potentially introduce some differences between computations and analytical solution for infinite size structure). Moreover, good agreement was achieved by manually fitting the values of edge-detecting filter that could lack accuracy for arbitrary structure different from overlapping disks. In other words, we still lack robustness in surface CFs evaluation due to inability to directly compare "discrete" versus "continuous" methodology, and verify both approaches for arbitrary geometries.

If computed accurately for any arbitrary structure at hand, surface CFs can be utilized in a multitude of applications:
\begin{enumerate}
  \item describing real surfaces resulting from different mechanical processes, e.g., cracking \cite{some_examples},
  \item delineating surface evolution \cite{chen2022} during such mechanical cracking events or reactive transport in porous media \cite{godinho2016,noiriel2021,prokhorov2022},
  \item representing surface features in super-resolution and stochastic reconstruction techniques \cite{chen2020super,janssens2020,karimpouli2022}.
\end{enumerate}
The examples above a just a few potential applications that, due to ability for surface CFs to establish themselves as universal interface/surface descriptors, highlight the necessity to establish robust techniques for $F_{ss}$ and $F_{sv}$ evaluation.

In this contribution we continue to build upon foundational work of Ma and Torquato \cite{ma2018SS}
and recent findings of Samarin et al. \cite{our_SS} to develop a robust and efficient approach to compute surface correlation functions from digital 2D and 3D images. In partiuclar, here our aim is to establish a direct verification of "discrete" methodology (that is easily parallelizable on CPU and GPU architectures) using exact noiseless "continuous" approach for arbitrary structures (still described by some continuous function). The rest of the manuscript is organized
as follows: in \cref{sec:def} we introduce some major definitions of CFs necessary for establishing algorithms, in \cref{sec:algo} we describe the all methodological details for "discrete" method for digital images. Next, in
\cref{sec:verify} we explain in detail why verification of computational approaches for digital images is currently hampered, and then introduce
novel exact "continuous" methodology in \cref{sec:cool-algo}. The choice of verification cases is presented in \cref{sec:fss-2d} followed by improvements in methodology of Samarin et al. The paper concludes with discussion, outline and summary in \cref{sec:summary}.

@@

Now we define two disjoint sets in a homogeneous porous medium: a set of solid
phase $S$ and a set of void phase $V$. Surface-surface correlation function
$F_{ss}$ and surface-void correlation function $F_{sv}$ are then defined as follows:
\begin{align}
  F_{ss}(\bm{r}) &= \langle |\nabla \chi_S(\bm{x})| |\nabla \chi_S(\bm{x} +
  \bm{r})| \rangle \label{eq:fss} \\
  F_{sv}(\bm{r}) &= \langle |\nabla \chi_S(\bm{x})| \chi_V(\bm{x} +
  \bm{r}) \rangle \label{eq:fsv}
\end{align}
Here the brackets $\langle \dots \rangle$ denote ensemble average. For isotropic homogeneous media
vector values $\bm{x}$ and $\bm{r}$ can be replaced with scalars. In case correlation functions are computed
in predefined directions, scalar values can be also used for anisotropic structures \cite{jiao2014chawla,EPL1}.

More detailed information on surface correlation functions can be found in comprehensive
Torquato's book \cite{Torquato_book}.

@@

From now on we will work with square images with dimensions $1 \times 1$ square
length units, so we can omit division by $N$ in described algorithms and
ensemble average in \cref{eq:fss} and \cref{eq:fsv} becomes a double integral
over a square $[0, 1]^2$.

@@

\subsection{Surface-surface function in two-dimensional case}
\label{sec:fss-2d}
\begin{figure}
  \centering
  \includegraphics[width=0.8\linewidth]{images/Fss.png}
  \caption[]{Illustration of how $F_{ss}(\bm{r})$ is calculated. The interface
    between solid and void phases is a solid black curve. This interface
    translated by the vector $\bm{r}$ is show as a red solid curve. Support of
    $f_\Gamma(\bm{x}; \Delta)$ is a dashed ring. Shaded regions contribute to
    the surface-surface function.}
  \label{fig:Fss-explained}
\end{figure}
Let $S$ be some set in $\mathbb{R}^2$ which has a smooth boundary $\Gamma$ in
the sense that the boundary has a tangent line almost everywhere. Now we
introduce a function $f_\Gamma(\bm{x}; \Delta)$:
\begin{equation*}
  f_\Gamma(\bm{x}; \Delta) = \left\{
  \begin{array}{ll}
    1/\Delta & \quad \rho(\bm{x}, \Gamma) < \Delta \\
    0 & \quad \text{otherwise}
  \end{array}
  \right.
\end{equation*}
Here $\rho(\bm{x}, \Gamma)$ means a closest distance from point $\bm{x}$ to the
boundary $\Gamma$.

@@

It's useful to introduce an imaginary unit $\varepsilon$ and write a dual number
$(x, y)$ as $x + y\varepsilon$. As you can see from multiplication law,
$1\cdot \varepsilon = \varepsilon$ and $\varepsilon^2 = 0$, hence $D$ is not an
integral domain and division is not defined \textcolor{red}{for all dual numbers with exception
of $0$.} Namely it's defined only if divisor has non-zero real part:
\begin{align*}
  \frac{a+b\varepsilon}{c+d\varepsilon} &=
  \frac{(a+b\varepsilon)(c-d\varepsilon)}{(c+d\varepsilon)(c-d\varepsilon)} \\
  &= \frac{ac-ad\varepsilon+bc\varepsilon}{c^2} \\
  &= \frac{a}{c} + \frac{bc-ad}{c^2}\varepsilon
\end{align*}

@@

\bibliography{paper}
